---
title: "Baseline Correction"
author: "Candace Savonen"
date: "2024-09-17"
output:
  pdf_document: default
  html_document: default
---

## Install hyperSpec

This will only install if it isn't already installed.

```{r}
if (!"hyperSpec" %in% installed.packages()) {
  install.packages("hyperSpec")
}
library(magrittr)
library(hyperSpec)
```

This will ask you to specify the file you are trying to baseline correct.

```{r}
if (interactive()) {
  data_file <- file.choose()
} else {
  data_file <- "/Users/jonathan.bennion/OneDrive - US Army/Documents/R/Files from Candy/DNDA-6_Pure_300 sec_10mW_center 2050 2024-05-15 12_37_26.csv"
}
```

This will read in the file you specified as is.

```{r}
raw_data <- readr::read_csv(data_file,
                            col_names = c("wavelength", "intensity"),
                            col_types = "nn") %>% 
  dplyr::filter(wavelength > 325) %>%
  dplyr::filter(wavelength < 3100)
```

Set up special object from the raw data.

```{r}
spec_obj <- new(
  "hyperSpec",
  spc = t(raw_data),
  wavelength = as.vector(raw_data$wavelength))


sqrt_obj <- sqrt(rowMeans(spec_obj))
```

Plot it before we correct to see.

```{r}
plotspc(spec_obj)
```

Find the baseline:

```{r}
baseline1 <- spc.rubberband(spec_obj, noise = 10, df = 9)

baseline2 <- spc.fit.poly.below(spec_obj, 
                               poly.order = 5,
                               debuglevel = 20, 
                               stop.on.increase = FALSE)
```

Subtract the baseline

```{r}
corrected_data1 <- spec_obj - baseline1
corrected_data2 <- spec_obj - baseline2
```

Plot it so we can take a look if its accurately corrected.

After correction:

```{r}
plotspc(corrected_data1)
```

```{r}
plotspc(corrected_data2)
```

```{r}
corrected_data1_df <- data.frame(wavelength = corrected_data1@wavelength,
                                 corrected_intensity = corrected_data1$spc[2,])
corrected_data2_df <- data.frame(wavelength = corrected_data2@wavelength,
                                 corrected_intensity = corrected_data1$spc[2,])

readr::write_csv(corrected_data1_df, gsub(".csv$", "_bl1_baseline.csv", data_file))
readr::write_csv(corrected_data2_df, gsub(".csv$", "_bl2_baseline.csv", data_file))
```


## Session Info
This prints out what packages and versions were being used for this:

```{r}
sessionInfo()
```
