---
title: "Raman Plots"
output:   
  html_notebook: 
    toc: true
    toc_float: true
author: Candace Savonen
date: 2020
---

## Set Up

This uses algorithms from `baselineWavelet`, so we need to install this if 
it hasn't been installed. 

```{r echo=FALSE}
if (!("tidyverse" %in% installed.packages())){
  install.packages("tidyverse")
}

if (!("baselineWavelet" %in% installed.packages())){
  package_file <- file.path("util", "zmzhang-baselineWavelet-4.0.2-9.zip")
  
  # Download the package
  download.file("https://github.com/zmzhang/baselineWavelet/archive/981c41095f5b4aaedf06bbcb2f4366acd467d1f0.zip", 
                destfile = package_file)
  
  install.packages(package_file, 
                   repos = NULL, type = "source")
}

if (!("patchwork" %in% installed.packages())){
  devtools::install_github("thomasp85/patchwork")
}
```

Attach libraries we need. 

```{r}
# magrittr pipe
`%>%` <- dplyr::`%>%`

library(baselineWavelet)
library(Matrix)
library(patchwork)
```

Import special functions. 

```{r}
source(file.path("util", "read_n_correct.R"))
source(file.path("util", "plotting_wrappers.R"))
```

### Directories and Files

The plots we make will be saved to this directory.

```{r}
# Path to output directory
plots_dir <- "plots"

# Create the plots_dir if it does not exist
if (!dir.exists(plots_dir)) {
  dir.create(plots_dir)
}

# Call the plot what you would like it to be saved as
plot_filename <- file.path(plots_dir, "<PLOT_NAME>")
```

Any files that are in this directory will be run. 

```{r}
# Path to input directory
input_dir <- file.path("<FOLDER_NAME>")

# Check that what you supplied exists
dir.exists(input_dir)
```

Get all the file names!

```{r}
# Get the list of files from here
input_files <- list.files(input_dir, 
                          pattern="^.*\\.(csv|CSV)$", 
                          full.names = TRUE)

# Test if all files found?
all(sapply(input_files, file.exists))
```

Collect experiment information from the file names.

```{r}
input_df <- data.frame(input_files) %>%
  # Split up file names into multiple columns so we can handle it easier
  tidyr::separate(input_files, sep = "_", 
                  into = c("file_path", 
                           "name", 
                           "pressure_pt", 
                           "pressure", 
                           "wavelength"
                           )) %>% 
  # Save input file paths as their own column too
  dplyr::mutate(input_files = input_files, 
                pressure = as.numeric(stringr::str_remove(pressure, " GPa"))) %>% 
  # Filter to only sets that have P (not dP)
  dplyr::filter(grepl("^P", pressure_pt)) %>%
  # Select only the columns we'll move forward with
  dplyr::select(input_files, 
                pressure_pt, 
                pressure, 
                wavelength
                )

input_df
```

Neaten up the wavelength column. 

```{r}
input_df <- input_df %>%
  # Make the wavelength column only have the wavelength -- drop the dates
  dplyr::mutate(wavelength = as.numeric(stringr::word(wavelength, sep = " ", 2))) %>% 
  # Order the data.frame by the wavelength 
  dplyr::arrange(pressure, wavelength)

input_df
```

Where there's a 2800 and a 3200 for an individual pressure, only keep the 3200 file. 

```{r}
#too_many <- input_df %>%
#  dplyr::group_by(pressure)%>%
#  dplyr::summarize(n = dplyr::n()) %>% 
#  dplyr::filter(n > 3) %>% 
#  dplyr::pull(pressure)

#input_df <- input_df %>% 
#  dplyr::filter(!(pressure == too_many & wavelength == 2800))
```

Print out the final version of the data frame with the input files. 

```{r}
input_df
```

## Specify the pressures and ranges for each file

Name the pressures for each file and make a table that explains the 
samples being plotted. 

**Double check the table printed out that it has the accurate pressure and range labels.**

```{r}
# Make a data.frame with each samples' information
sample_df <- input_df %>%
  dplyr::mutate(wave_min = dplyr::recode(wavelength, 
                          "375" = 290, 
                          "2275" = 1400, 
                          "2800" = 2800),
                wave_max = dplyr::recode(wavelength, 
                          "375" = 1310, 
                          "2275" = 1875, 
                          "2800" = 3400) 
                          ) 

# Print out the table so it can be double checked
knitr::kable(sample_df %>% dplyr::select(pressure, wavelength, wave_min, wave_max))
```

## Run each data file through the correction algorithms

Use the specialized function to import and correct the data from each file.

```{r}
# Read, background correct, and identify peaks of each sample
corrected_datasets <- apply(sample_df, 1, 
                            function(file) {
                            read_n_correct(filename = file['input_files'], 
                                           wavenum_min = as.numeric(file['wave_min']),
                                           wavenum_max = as.numeric(file['wave_max']))
                              })

# Bring along the sample labels
names(corrected_datasets) <- paste0(sample_df$pressure, 
                            "GPa",
                            sample_df$wave_min,
                            "-", 
                            sample_df$wave_max,
                            each = "")
```

## Obtain a individual plot for each corrected dataset

```{r}
peak_plots <- lapply(corrected_datasets, 
                     plot_ind_peaks)

names(peak_plots) <- paste0(sample_df$pressure, 
                            "GPa",
                            sample_df$wave_min,
                            "-", 
                            sample_df$wave_max,
                            each = "")
```

## Assemble the main plot

Put together all the peak plots into one grid. 
This custom function assembles the individual plots where each row is a pressure. 

```{r}
ultimate_plot <- 
  assemble_main_plot(peak_plots_list = peak_plots,
                     row_pressures = input_df$pressure,
                     main_title = "The background-correction of Raman Spectra",
                     x_lab = "Wavelength / nm",
                     y_lab = "Raman Intensity/Arbitr. Units")
```

Print out preview of the final plot. 

```{r}
ultimate_plot
```

Save plot to a png

```{r}
ggplot2::ggsave(
  filename = file.path(plot_filename), 
  plot = ultimate_plot, 
  width = 10, 
  height = 12)
```

## Session Info

```{r}
sessionInfo()
```

