---
title: "Raman Plots"
output:   
  html_notebook: 
    toc: true
    toc_float: true
---

## Set Up

This uses algorithms from `baselineWavelet`, so we need to install this if 
it hasn't been installed. 

```{r}
if (!("baselineWavelet" %in% installed.packages())){
  httr::set_config( httr::config( ssl_verifypeer = 0L ) )
  devtools::install_github("zmzhang/baselineWavelet")
}
# Magrittr pipe
`%>%` <- dplyr::`%>%`
```

### Directories and Files

The plots we make will be saved to this directory.

```{r}
# Path to output directory
plots_dir <- "plots"

# Create the plots_dir if it does not exist
if (!dir.exists(plots_dir)) {
  dir.create(plots_dir)
}

# Call the plot what you would like it to be saved as
plot_filename <- "raman_plot.png"
```

Any files that are in this directory will be run. 

```{r}
# Path to input directory
input_dir <- "input_data"

# Get the list of files from here
input_files <- list.files(input_dir, 
                    pattern="^.*\\.(csv|CSV)$", full.names = TRUE)

# Print out file names in this directory
cat(
  "The following files will be run: \n",
  paste(input_files, collapse = "\n")
)
```

## Specify the pressures and ranges for each file

Name the pressures for each file and make a table that explains the 
samples being plotted. 

**Double check the table printed out that it has the accurate pressure and range labels.**

```{r}
# Make a data.frame with each samples' information
sample_df <- data.frame(input_files, 
                        pressure = rep(c("20GPa", "40GPa"), 3), 
                        range = factor(c("high", "high", "low", "low", "mid", "mid"), 
                                       levels = c("low", "mid", "high")),
                        stringsAsFactors = FALSE
                        ) %>% 
  dplyr::mutate(wave_min = dplyr::recode(range, 
                          "low" = 200, 
                          "mid" = 1500, 
                          "high" = 3000
                        ),
                wave_max = dplyr::recode(range, 
                          "low" = 1500, 
                          "mid" = 1800, 
                          "high" = 3500)
                ) %>%
  dplyr::arrange(pressure, range)

# Print out the table so it can be double checked
knitr::kable(sample_df)
```

## Set up special functions

```{r}
read_n_correct <- function(filename, wavenum_min = 0, wavenum_max = 4000) {
  # Will background correct and identify peaks for a given vibration dataset 
  # saved as a CSV file from
  #
  # Args:
  #   filename : A file path to to a vibration CSV file that has two columns
  #             The first column is wavenumber and the second is intensity.
  #   wavenum_min/max : two numbers which specify the range of numbers to keep in 
  #                    the dataset. Wavenumbers outside these ranges will be 
  #                    removed from the data.frame
  #
  # Returns:
  # A data.frame with background corrected intensities and peaks identified. 
  # Get file.path
  # 
  filename <- file.path(filename)
  
  # Check if the file actually exists
  if(!file.exists(filename)){
    stop(paste("No file at the given file path."))
  }
  
  # Read in the CSV data file colnames will be ignored
  vibration_df <- readr::read_csv(filename, col_names = FALSE) %>% 
    # Save the sensible column names
    dplyr::rename("wavenumber" = X1, 
                  "intensity" = X2) 
  # Establish the scale
  scales <- seq(1, 120, 1)
  
  # Find Continuous Wavelet Transform
  w_coefs <-  baselineWavelet::cwt(vibration_df$intensity,
                                   # The scales at which to perform CWT
                                   scales = scales, 
                                   # Which wavelet base to use
                                   wavelet = 'mexh')
  # Make this dat into an image
	print(
	  image(1:nrow(w_coefs), 
	        scales,
	        w_coefs, 
	        col = rainbow(256), 
	        axes = FALSE, 
	        xlab = 'index', 
	        ylab = 'CWT coefficient scale', 
	        main = 'CWT coefficients')
	)

	# Identify the local maximum of each column in 2-D CWT coefficients matrix
	# by using a slide window
	local_max <- baselineWavelet::getLocalMaximumCWT(w_coefs, 
	                                                # Minimum window 
	                                                minWinSize = 0.5)
  # Identify ridges by connecting the local maximum of 2-D 
	ridge_list <- baselineWavelet::getRidge(local_max,
	                                        gapTh = 3, 
	                                        skip = NULL)
	
	# Identify the major peaks
	major_peak_info <- baselineWavelet::identifyMajorPeaks(vibration_df$intensity,
	                                                       ridge_list,
	                                                       w_coefs,
	                                                       SNR.Th = 0.2,
	                                                       ridgeLength = 1,
	                                                       nearbyPeak = TRUE)
	# Estimate peak width
	peak_width <- baselineWavelet::widthEstimationCWT(vibration_df$intensity,
	                                                  major_peak_info)
	# Plot the ridge list
	baselineWavelet::plotRidgeList(ridge_list)
	
	# Make the data.frame with the corrected data columns. 
	vibration_df <- vibration_df %>% 
	  dplyr::mutate(
	    # Make background a new column based on intensity column
	    background = baselineWavelet::baselineCorrectionCWT(
	      intensity,
	      peak_width,
	      lambda = 1000,
	      differences = 1),
	    # Save corrected intensity as a new column 
	    corrected_intensity = intensity - background, 
	    # Save binned column
	    binned = lowess(wavenumber, 
	                    corrected_intensity, 
	                    f = 0.03)$x, 
	    # Save uncorrected bin column
	    uncorrected_bin = lowess(wavenumber, 
	                             intensity, 
	                             f = 0.03)$x, 
	    # Save new column that says which are major peaks
	    major_peak = (1:nrow(.) %in% major_peak_info$peakIndex)
	    ) %>% 
	  dplyr::filter(wavenumber > wavenum_min, 
	                wavenumber < wavenum_max) 
	
	# Return the data.frame with the corrected columns
	return(vibration_df)
}
```

Function for plotting the individual peaks

```{r}
plot_ind_peaks <- function(vibration_df) {
  # Given a vibration data.frame that has been corrected already using 
  # `read_n_correct` function, make an individual peak shape for ultimate plot
  #
  # Args:
  #   vibration_df : A corrected vibration dataset as a data.frame with 
  #                  `corrected_intensity`, `wavenumber` and `major_peak` as 
  #                  columns. 
  #
  # Returns:
  # A data.frame with background corrected intensities. 
  
  # Make a plot with no axises or other stuff, but has the peaks
	ggplot2::ggplot(vibration_df) + 
	  ggplot2::geom_line(ggplot2::aes(x = wavenumber, 
	                                  y = corrected_intensity)) +
	  ggplot2::geom_point(data = subset(vibration_df, major_peak), 
	                      ggplot2::aes(x = wavenumber, 
	                                   y = corrected_intensity, 
	                                   shape = major_peak)) +
	  ggplot2::theme_classic() + 
    # Make everything blank because we just want the peak shape
    ggplot2::theme(legend.position = "none", 
                   axis.ticks.y = ggplot2::element_blank(),
                   axis.text.y = ggplot2::element_blank(),
                   axis.title.x = ggplot2::element_blank(),
                   axis.title.y = ggplot2::element_blank(), 
                   axis.line = ggplot2::element_blank()) 
}
```

## Run each data file through the correction algorithms

Use the specialized function to import and correct the data from each file.

```{r}
# Read, background correct, and identify peaks of each sample
corrected_datasets <- apply(sample_df, 1, 
                            function(file) {
                            read_n_correct(filename = file['input_files'], 
                                           wavenum_min = file['wave_min'],
                                           wavenum_max = file['wave_max'])
                              })

# Bring along the sample labels
names(corrected_datasets) <- paste0(sample_df$pressure, sample_df$range, each = "")
```

## Obtain a individual plot for each corrected dataset

```{r}
peak_plots <- lapply(corrected_datasets, 
                     plot_ind_peaks)

names(peak_plots) <- paste0(sample_df$pressure, sample_df$range, each = "")
```

## Assemble the main plot

Put together all the peak plots into one grid. 

```{r}
ultimate_plot <- cowplot::plot_grid(plotlist = peak_plots,
          align = "hv", 
          axis = "lbt",
          label_x = "Wavelength / nm", 
          label_y = "Raman Intensity/Arbitr. Units")
```

Declare x and y labels and main title. 

```{r}
# Declare the x label
x_lab <- grid::textGrob("Wavelength / nm")

# Declare the y label
y_lab <- grid::textGrob("Raman Intensity/Arbitr. Units", rot = 90)

# Declare the main title
main_title <- grid::textGrob("The background-correction of Raman Spectra")
```

Add these labels to the main plot!

```{r}
ultimate_plot <- gridExtra::grid.arrange(
  gridExtra::arrangeGrob(ultimate_plot, 
                         left = y_lab, 
                         bottom = x_lab, 
                         top = main_title)
  )
```

Save plot to a png

```{r}
cowplot::save_plot(filename = file.path(plots_dir, plot_filename), 
                   plot = ultimate_plot)
```

## Session Info

```{r}
sessionInfo()
```

